%!TEX root = JakubJedryszek-MasterThesis.tex
% +--------------------------------------------------------------------+
% | Abstract Page
% +--------------------------------------------------------------------+

\pagestyle{empty}
%\vspace{1cm}
\setlength{\baselineskip}{0.8cm}

\indent

% +--------------------------------------------------------------------+
% | Enter the text of your abstract below, maximum of 350 words.
% +--------------------------------------------------------------------+

The future of Medical Devices is their interoperability. Nowadays, medical devices works rather independently, which cause accidents we could avoid if different devices would communicate. Dr. Julian Goldman developed idea of "Integrated Clinical Environment" (ICE). It is series of standards, which describes medical device interoperability. SAnToS lab created Medical Device Coordination Framework (MDCF), which is prototype implementation of ICE idea.

AADL (Architecture Analysis \& Design Language) is modeling language for representing hardware and software. It is used for real-time, safety critical and embedded systems. AADL allows for the description of both software and hardware parts of a system. It is used to describe architecture, but AADL allows to add behavioral extensions through annex languages. BLESS (Behavior Language for Embedded Systems with Software) is AADL annex sub language defining behavior of components. The goal of BLESS is automatically-checked correctness proofs of AADL models of embedded electronic systems with software.

Ada is one of the most popular (along with C/C++) programming language targeted at embedded and real-time systems. SPARK Ada is subset of Ada, designed for the development of safety and security critical systems. It contains subset, which allows to reason about and prove correctness of program and its entities.

Nowadays, there is a trend to generate code from models. The ultimate goal of research, which this thesis if part of, is to create AADL/BLESS to SPARK Ada translator. Ultimately set of standardized AADL/BLESS models for medical devices would be created. From these models SPARK Ada code base would be generated. It would be further extended by developers.

This thesis propose mapping from AADL/BLESS to SPARK Ada. As an example of Medical Device, PCA (Patient Controlled Analgesia) pump is used. The foundation for this work is "Integrated Clinical Environment Patient-Controlled Analgesia Infusion Pump System Requirements" document \cite{PcaReq} and AADL Models with BLESS annexes created by Brian Larson. In addition to proposed mapping, PCA pump prototype was created. As a platform for prototyping, BeagleBoard-xM device was used. Some components of PCA pump prototype are verified by SPARK tools and Bakar Kiasan.
