% +--------------------------------------------------------------------+
% | Acknowledgements Page (Optional)                                   |
% +--------------------------------------------------------------------+

\newpage
\vspace*{0.9cm}
\begin{center}
{\bf \Huge Acknowledgments}
\end{center}

\setlength{\baselineskip}{0.8cm}

%\pdfbookmark[0]{Acknowledgements}{PDF_Acknowledgements}

% +--------------------------------------------------------------------+
% | Enter text for your acknowledgements in the space below this box.  |
% |                                                                    |
% +--------------------------------------------------------------------+

//Stolen from Venkatesh

To be involved in a research project is a great experience.

I am thankful to Dr.Matthew Dwyer for offering me an opportunity (which I did grab!) to work in BANDERA group. I am thankful to Dr.John Hatcliff, my major professor, for being patient and persistent while guiding and mentoring me while working on this thesis. I am thankful for the numerous opportunities they both gave me to be involved in activities other than my thesis, and most of all for being enthusiastic and supportive. I am also thankful to Dr.Anindya Banerjee for providing useful inputs during the preparation of the thesis.

I am thankful to the entire BANDERA group for providing a great working envi- ronment. In particular, I am thankful to Hongjun Zheng for being a rigorous tester of my implementation. I am thankful to my friends for bearing with my antics, and making the long hot summers and bitter cold winters bearable.

I am thankful to my wife, Prathima, for being patient and supportive during my indulgence with my workstation for long hours and for bearing with my rather erratic feeding habit.

They say, save the best for last. Hence, lastly I would like to thank my mother, my father, and my elder sister for advising me in the ways of life and helping hold the helm during storms. “Home is the first school” is an adage in Kannada. I am indebted to my family for providing a great first school.
