%!TEX root = JakubJedryszek2014.tex

\cleardoublepage

\chapter{Future Work}
\label{future_work}

\begin{chapquote}{\textit{Benjamin Franklin}}
``If you fail to plan, you plan to fail.''
\end{chapquote}

The following are possible extensions for work done in this thesis:

\begin{itemize}
	\item The most important thing, which would be extremely helpful to proceed with work done in this thesis, would be to review it by some industry expert and experienced engineer. Especially, how particular functionalities (like monitoring external sensors or controlling pump actuator) are implemented, and how looks communication between components modeled in AADL.
	\item Creation of automatic translator described in Section \ref{codegen:translator} would be good validation of created translation schemes. It may reveal some issues not present for manual translation.
	\item Currently AADL thread properties are not taken into account in thread to task mapping, in Section \ref{codegen:mapping:threads}. Properties like priority or period could be mapped pretty straightforward to SPARK Ada. For now, former is hard-coded as 10 and latter simply skipped, which requires developer to handle it. However, given properties that are modeled and analyzed in AADL models, should be translated automatically to maintain synchronization between models and the code. AADL properties are described in \cite{AadlBook}, in the Appendix A.
	\item Data types translation presented in Section \ref{codegen:mapping:data}, in addition to straightforward type mapping, includes protected types. However, all protected types have the same set of subprograms (\lstinline{Put} and \lstinline{Get}). It is worth to consider introduction of generics, which will allow to specify generic protected type and then reuse it for all types.
	\item In feature groups translation (Section \ref{codegen:mapping:feature_groups}), idea of child or nested packages was abandoned. However, it would be good to reconsider it for the purposeof providing encapsulation for grouped features. It may be useful to introduce getter functions in parent package or some other technique that would allow for better separation and decomposition.
	\item AADL property set mapping in Section \ref{codegen:mapping:propertyset} handles only \lstinline{aadlinteger} type. Thus, it requires extension for handling other more complex constructs.
	\item Current translation schemes cause creation of pretty big packages, which will become bigger after adding implementation. Thus, some decomposition is desired. The following techniques can be considered: 
		\begin{itemize}
			\item partitioning of packages
			\item taking advantage of child packages
			\item separation of threads to different packages (e.g. one thread per child package and all common functionalities in parent package)
			\item simple package separation
		\end{itemize} 
	\item The mappings for BLESS are limited only to a small subset. Development of translations for BLESS state machines (states and transitions) would be good addition. this would support behavior translation. A good point to start is the \lstinline{Rate_Controller} thread, which can be found in \lstinline{PCA_Operation_Threads} package in original AADL models created by Brian Larson. The semantics of BLESS contain notions of time that make translation to SPARK difficult. This problem occurs in state machine models. Finding solutions for that is needed. Maybe even, by changing BLESS semantics.
	\item When this thesis was written, SPARK 2014 did not support multitasking. However, there were plans to introduce it into SPARK 2014 following an approach similar to SPARK 2005. Once multitasking support is present, translations for SPARK 2014 will be possible.
	\item There is an issue with two way communication between SPARK packages caused by circular dependency. It is described in Section \ref{codegen:port_communication:thread}.
	\item The port communication presented in Section \ref{codegen:port_communication} captures only 1:1 connections between ports of the same type and opposite direction. In AADL there are also inter-port connections and one-to-many or many-to-one connections. \cite{AadlBook} They should be taken into AADL subset for medical devices modeling and translation. 
	\item The created PCA pump prototype contains only basic functionalities. Some parameters (like drug concentration) are ignored. The next step is its development should be to include funcionality that has been omitted. In addition to that, interaction with external modules such as sensors for monitoring drug flow or communication with ICE through Ethernet port is desired. This requires creation of communication channel between BeagleBoard (SPARK Ada application) and these systems. 		
\end{itemize}
