%!TEX root = JakubJedryszek-MasterThesis.tex

\cleardoublepage

\chapter{Summary}
\label{summary}

In this thesis PCA Pump prototype, in SPARK 2005 with Ravenscar profile, has been created. It runs on BeagleBoard-xM platform and control physical device. Furthermore, AADL/BLESS to SPARK Ada translation is proposed. Based on that sample translation from simplified AADL models of PCA pump has been performed. At the end, example verification of created prototype, isolated module for dose monitoring and translated code has been shown. 

All work done in this thesis targets SPARK 2005. SPARK 2014 and its tools (such as GNATprove) were not ready at the time, when this thesis was written. However, an example verification (of dose monitoring module, which has been translated to SPARK 2014) was presented.

% lack of resources
% no access to real-live examples
The biggest challenge during PCA pump development was the SPARK limitations. There are many common libraries, which cannot be verified by SPARK tools. Thus it is required to isolate some functionalities or implement them in different way. Another issue was lack of resources and SPARK code samples. Especially industry code examples, which are keep secretly as intellectual property by companies. Available resources are usually small examples used in research or reference manuals, which were created couple years age. Although still valid, not updated for years.

% everything 'under development'
Furthermore, BLESS and SPARK are still under development. Thus, it was very hard to take advantage of all desirable capabilities (most of features are not yet implemented). An example may be lack of support for pre- and post conditions in RavenSPARK.

% small community
In addition to that, community working with above technologies is very small. On StackOverflow\footnote{http://stackoverflow.com} there is 728 question related to Ada\footnote{http://stackoverflow.com/questions/tagged/ada} and only 3 to SPARK\footnote{http://stackoverflow.com/questions/tagged/spark-ada}. In the same time, C\# has 673,721 questions\footnote{http://stackoverflow.com/questions/tagged/c\%23} and Java - 682,308\footnote{http://stackoverflow.com/questions/tagged/java}.

% proposed mapping is probably wrong
% no consultancy with industry expert/programmer
Proposed mapping from AADL to SPARK Ada is not consulted with industry engineers. Thus, it would be first thing to do, in order to continue this research. Lot of work can be done in this topic. It is described in chapter \ref{future_work}.